---
title: 'Anomaly (Outlier) Detection'
author: "Illarion  Jabine"
date: "12/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


### Required packages:

* [AnomalyDetection]: Anomaly Detection Using Seasonal Hybrid Extreme Studentized Deviate Test
* [outliers]: Tests for outliers 
* [devtools]: Required to install AnomalyDetection and isofor from github
* [FNN]: Fast Nearest Neighbor Search Algorithms
* [dbscan]: Density Based Clustering of Applications with Noise
* [isofor]: Isolation Forest Anomaly Detection


### Key terms
 * Grubbs' test 
 * Seasonal Hybrid ESD algorithm
 * Distance and density based anomaly detection
 * kNN distance matrix
 * kNN distance score
 * Density based clustering
 * Local Outlier Factors (LOF)
 * Isolation Forest anomaly detection
 * Grid
 * Anomaly contours


### Useful Links
<https://github.com/twitter/AnomalyDetection>
<https://github.com/Zelazny7/isofor>

## Introduction

Anomaly or outlier detection is the process of assessing whether data contain unusual points. Points that stand out "away" from the major hord of points.
A good definition of an anomaly:
It is a data point or collection of data points that do not follow the same pattern 
or have the same structure as the rest of the data.
Real life examples can be a sudden spike in a normal flow of credit card transactions,
or bizzare event in normal behaviour.
 * A point anomaly is a single data point which is unusual when compared to the rest of the data.
 * Collective anomaly is an anomalous collection of data points. They are unusual when considred together.
The anomaly detection is widely used in fraud, intrusion etc detection problems.
In this manual I will cover anomaly detection for:
 1. Univariate data
 2. Time series
 3. Multivariate data.

### 1. Load the libraries
Let's first load the libraries.
Note: When installing isofor and AnomalyDetection, the system will ask you to install RBuildTools. 
Click Yes, wait once it's installed, restart R and relaunch the package installation.
```{r loading packages, message=FALSE, warning=FALSE}
devtools::install_github("twitter/AnomalyDetection")
devtools::install_github("Zelazny7/isofor")
library(AnomalyDetection)
library(isofor)
library(outliers)
library(FNN)
library(dbscan)

```

### 2. Loading and checking the data

Load and check the datasets.
River dataset:
* index - the order of the nitrate observations 
* nitrate - monthly concentrations of dissolved nitrate found in a river
* month - a factor containing the month for each nitrate observation
Wine dataset: different characteristics of wines.
```{r load the data and pre-process them}
# Loading data from Rds file
load("data/anomaly_detection.Rds")

# Checking if there are any NAs:
anyNA(river_nitrate)
anyNA(wine)

```

### 3. Univariate anomaly test: Grubbs' test

Before using Grubbs' test, we have to check if the data are normally distributed.
hist() function can help to judge about the normal assumption.
```{r Grubbs' test}
attach(river_nitrate)
hist(nitrate, xlab = "Nitrate concentration", breaks = 30)
boxplot(nitrate, ylab = "Nitrate concentration")
summary(nitrate)
# Now let's apply Grubbs' outlier test
grubbs.test(nitrate)

```

The logic of the Grubbs' test is the following:
the lower the p-value returned by the test, the higher the likelihood that the point tested was an outlier.
It looks like the maximum value in nitrate variable is indeed outlier.
Let's find its index delete this value and run test again.

```{r another outlier?}
grubbs.test(nitrate[-which.max(nitrate)])
# p-value > 0.05, so the next point is not outlier.

# However, boxplot still treats it as an outlier.
boxplot(nitrate[-which.max(nitrate)])
```

### 4. Anomaly Detection in a Time Series

Great article on anomaly detection and AnomalyDetection package:
<https://blog.twitter.com/engineering/en_us/a/2015/introducing-practical-and-robust-anomaly-detection-in-a-time-series.html>


Let us first do explorative study of the time series
```{r exploring time series}
# Show the time series of nitrate concentrations with time
plot(nitrate ~ index, type = "o")

# Calculate the mean nitrate by month, using tapply(): tapply(X, INDEX, FUN = NULL):
# X object over which we apply function
# INDEX a list of factors used for splitting.
monthly_mean <- tapply(nitrate, months, FUN = mean)

# Plot the monthly means 
plot(monthly_mean, type = "o", xlab = "Month", ylab = "Monthly mean")


# Create aset of boxplots of nitrate against months
boxplot(nitrate ~ months)

```
From the initial analysis we could conclude that ther is a repeating seasonal cycle of 1é months.
Let's now use Seasonal-Hybrid ESD algorithm to understand where the anomalies occur within the time series data.

### 4.1 Seasonal-Hybrid ESD algorithm

AnomalyDetectionVec() function from AnomalyDetection package generates a list with the two elements:
 *anoms: Data frame containing index, values, and optionally expected values.
 *plot: A graphical object if plotting was requested by the user. The plot contains the estimated anomalies annotated on the input time series.

```{r SH ESD}
anomalies_sh_esd <- AnomalyDetectionVec(x = river_nitrate$nitrate, period = 12, direction = 'both', plot = T)

# Print the anomalies
anomalies_sh_esd$anoms

# Print the plot
print(anomalies_sh_esd$plot)
```

### 5. Distance and Density Based Anomaly Detection

Here I will use two functions:
 1. FNN::get.knn() - distance based anomaly detection
 2. dbscan::lof() - Local outlier factor score
k-nearest neighbors distance and local outlier factor use the distance or relative density of the nearest neighbors to score each point.
### 5.1 Distance based anomaly detection

FNN::get.knn() creates a kNN distance matrix.
This matrix has n rows, where n is the number of data points 
and k  columns, where k is the user-chosen number of neighbors.
The entry in row i and column j of the distance matrix is the distance between point i and its jth nearest neighbor.
The distance matrix is only appropriate for numeric continuous data.
Another important consideration is standardizing (with similar means and quartiles) features. Without standardization, features have different scales.

```{r kNN matrix}
anomalies_knn <- get.knn(scale(wine), k = 10)

# View the distance matrix
head(anomalies_knn$nn.dist)
```

Now let's calculate the nearest neighbor distance score.
It can be calculated by averaging the nearest neighbor distances for each point. 
Large values of the distance score can be interpreted as indicating the presence of unusual or anomalous points.
```{r the nearest neighbor distance score}
# Create score by averaging distances
knn_distance_scores <- rowMeans(anomalies_knn$nn.dist)

# Append the score as a new column 
wine$score_knn <- knn_distance_scores

# Let us plot the sulphates, alcohol and score_knn:
plot(sulphates ~ alcohol,data = wine, cex = sqrt(score_knn), pch = 20)

```

### 5.2 Local Outlier Factors (LOF) using dbscan

It's said that kNN is useful for finding global anomalies, whereas for local outliers we can use LOF.
I will use lof() function from dbscan package.

```{r LOF}
# k: the number of neighbors used to calculate the LOF.
lof_score <- lof(scale(wine[-14]), k = 10)

# Append the LOF score as a new column
wine$score_lof <- lof_score

# Scatterplot showing the sulphates, alcohol and LOF score
plot(sulphates ~ alcohol, data = wine, cex = sqrt(score_lof), pch = 20)

```

### 5.3 kNN vs LOF

```{r}
#Print the row indices of the wine data with highest kNN score and highest LOF score.
# Find the row location of highest kNN
which.max(wine$score_knn)

# Find the row location of highest LOF
which.max(wine$score_lof)

# Local anomalies aren't always the farthest from the rest of the data. 
```

### 6. Isolation forest algorithm

Isolation forest is a tree-based approach, a fast and robust method that measures how easily points can be separated by randomly splitting the data into smaller and smaller regions.
It can handl both numerical and categorical data.
I will use iForest() from isofor package.

Let's first build one single tree (nt = 1)
```{r single tree}
# let's exclude scores from knn and lof
wine_iso <- wine[-c(14,15)]

wine_1_iso_tree <- iForest(wine_iso, nt = 1)
```

